---
title: "Football analysis"
output: html_document
---
```{r}
# Libraries used
library(dplyr)
library(tidyverse)
library(tidyr)
```



## Prepare the data for analysis
### Read in data
```{r}
# #install package 'xlsx' if not installed yet
# if (!"xlsx" %in% installed.packages()) install.packages('xlsx')
# library('xlsx')
# 
# #read in data sheets
# sheet.index <- 1:3
# data.list <- list()
# for(i in sheet.index)
#   data.list[[i]]<-read.xlsx('football_salaries.xls', sheetIndex = i)
# 
# # Rename the datasets as we wanted
# player_salary <- data.list[[1]]
# coach_salary <- data.list[[2]]
# coach_record <- data.list[[3]]
```
Read from different files
```{r}
player_salary <- read.csv('player_salary.csv',
                    header = TRUE,
                    sep = ',')
coach_salary <- read.csv('coach_salary.csv',
                    header = TRUE,
                    sep = ',')
coach_record <- read.csv('coach_record.csv',
                    header = TRUE,
                    sep = ',')
```

### Cleaning data
#### General information of the datasets
We first take a brief look at the datasets of their descriptive statistics.
```{r}
summary(player_salary)
summary(coach_salary)
summary(coach_record)
```
Check if there are Na values.
```{r}
# Check if there are NA values
for (i in sheet.index){
  print(data.list[[i]] %>% is.na() %>% sum())
}
```
We see there are some cleaning should be done before we can use these datasets:  
1. Take care of the missing value in **coach_salaries**  
2. Change the column names in **coach_record**  
3. For consistency, change salary-related variable types to double in **coach_salary** and **coach_record**.    
4. Separate variables if needed.  

##### 1. missing values
```{r}
# replicate the original model for cleaning
player_salary_ <- player_salary
coach_salary_ <- coach_salary
coach_record_ <- coach_record

# Finding the missing value, and filled in with internet searching result
coach_salary_ <- coach_salary
coach_salary_[c(which(is.na(coach_salary)==TRUE, arr.ind = TRUE)[,1]),5] <- '2.5 million'
```

##### 2 Fix column names
```{r}
colnames(coach_record_) <- c(coach_record_[1,])
coach_record_ <- coach_record_[-1,]
rownames(coach_record_) <- c(1:nrow(coach_record_))

colnames(coach_salary_) <- c(colnames(coach_salary_)[1:3], 'Annual_salary', 'Net_worth', colnames(coach_salary_)[6])
```

##### 3 change the variable type   
3.1. **coach_salary**
* *Annual_salary*
```{r}
# get the index of the cells with out 'million'
no_mill <- c(which(!agrepl('million', coach_salary_$Annual_salary, ignore.case = T)))

# paste 'million' to uniform these cells
for (i in no_mill)
  coach_salary_[i,'Annual_salary'] <- paste(coach_salary_[i,'Annual_salary'], 'million', collapse = ' ')

# separate the numeric parts and drop million
coach_salary_ <- separate(coach_salary_, Annual_salary, c('income', NA), ' ')

# Keep numeric parts and change variable type to numeric
for (i in 1:nrow(coach_salary_))
  coach_salary_[i,'income'] <-
substr(coach_salary_[i, 'income'], 2, nchar(coach_salary_[i,'income']))

# numerize the variable
for (i in 1:nrow(coach_salary_)){
  if (i %in% no_mill)
    coach_salary_[i, 'income'] <- as.double(gsub(',', '', coach_salary_[i,'income']))
  if (i==24)
      coach_salary_[i, 'income'] <- as.double(substr(coach_salary_[i, 'income'], 1, (nchar(coach_salary_[i,'income'])-1)))
  else
    coach_salary_[i, 'income'] <- as.double(coach_salary_[i, 'income'])*1000000
}
coach_salary_ <- rename(coach_salary_, 'Annual_salary'='income')
```

* *Net_salary*
```{r warning=FALSE}
# Remove dollar sign and trim
for (i in 1:nrow(coach_salary_)){
  coach_salary_[i, 'Net_worth'] <- gsub('\\$', '', coach_salary_[i, 'Net_worth'])
  coach_salary_[i, 'Net_worth'] <- gsub(' ' , '', coach_salary_[i, 'Net_worth'])
}

coach_salary_ <- separate(coach_salary_, Net_worth, c('net', NA), 'million')

coach_salary_$net <- as.double(coach_salary_$net)*1000000

coach_salary_ <- rename(coach_salary_, 'Net_worth'='net')
```

3.2 **coach_record**
```{r}
for (i in 4:ncol(coach_record_))
  coach_record_[,i] <- as.double(coach_record_[,i])
```

##### 4. Separate variables  
* **player_salary**
```{r warning=FALSE}

player_salary_ <- player_salary %>% separate(free_agency, c('agency_time', 'type'),' ')
# There's warning of NA's in the dataset

# Extract rows with NA values
player_salary_[c(which(is.na(player_salary_)==TRUE, arr.ind = TRUE)[,1]),]

# These are not top players
# the number of missing data is not very large, so we just drop  them
player_salary_ <- player_salary_[-c(which(is.na(player_salary_)==TRUE, arr.ind = TRUE)[,1]) ,]
rownames(player_salary_) <- c(1:nrow(player_salary_))

player_salary_$agency_time <- as.numeric(player_salary_$agency_time)
```
#### 5, Fiex abnormal data
Replace age larger than 50 or smaller than 15 with mean age of the rest age data.
```{r}
age_mean <- player_salary_ %>% filter(age>15) %>% filter(age<50) %>% select(age)%>% summarize_each(funs(mean))

player_salary_[player_salary_$age < 15, 'age'] <- age_mean
player_salary_[player_salary_$age > 50, 'age'] <- age_mean
```
There are three 0 values in *avg_year*, so we just delete them.
```{r}
player_salary_ <- player_salary_[player_salary_$avg_year != 0,]
```

```{r}
player_salary <- player_salary_
coach_record <- coach_record_
coach_salary <- coach_salary_
```

## Topics and Analysis

# 5. How many players will be free agency by 2022?


```{r}
# Importing package
library(readxl)
# Loading dataframe
df_salaries <- read_excel("football_salaries.xls")
# Removing strings from column 
#gsub or str_replace
```
# 6. What are the 20 MVP players based on their earns?
```{r}
# Creating column with MVP value 
df_MVP <- df_salaries %>% 
  rowwise() %>% 
  mutate(MVP_value = sum(avg_year, total_guaranteed, fully_guaranteed))
# Removing duplicated players'name
df_MVP <- df_MVP %>% distinct(player, .keep_all= TRUE)
# Organizing it by sorting as ascending values
df_MVP <- df_MVP[order(df_MVP$MVP_value),]
# Extracting the 20 highest MVP_values
df_20 <- tail(df_MVP, n= 20)
# The 20 most valuable players (MVP) are:
df_20$player
# Which position has more MVP's?
df_20$position <- as.factor(df_20$position) # making position as factor
ggplot(df_20, aes(x= position, color= position)) +
    geom_bar(stat = "count") +
    labs(x = "Position") +
  ggtitle("MVP's per position")
```
The position that shows more MVP's is the quarterback.

# 7. What teams have the highest fully guaranteed expenses?

# 8. What is the relationship between coach salaries and their records?



### What are the variables related to high salary of players?
#### position
##### Average highest
```{r}
# unique(player_salary$position)#29 positions in the data

# Select columns of position, avg_year, player to find positions with the highest salary

position_sub<- player_salary[c('position','avg_year')]%>%
  group_by(position)

p <- position_sub %>%
      summarize_each(funs(mean)) %>%
      arrange(desc(avg_year)) %>%
      head(10) %>%
      select(position) %>%
      inner_join(position_sub, by='position') %>% 
      group_by(position) %>%
      ggplot(aes(x=position, y=avg_year))+
      geom_boxplot(aes(fill=avg_year)) +
      theme(axis.text.x = element_text(angle=90, hjust = 1))
p + stat_summary(fun=mean, color='darkred', geom='point') #red dot identifies the mean value
```
Though the boxplot quartiles are minimized by the outliers, the position with the highest yearly pay is **quarterback**. To clear see the tendency, zoom in to see the average values.


```{r}
p + stat_summary(fun=mean, color='darkred', geom='point') +
    coord_trans(x = "identity", y = "identity", xlim = NULL, ylim = c(0,6e+6))

```
##### Top highest in each position
We have known that the highest average is **quarterback**, but when we look at medians, they are not as much different as averages. The "outliers" make this discrepancy, so I would take a look at the top 10 ouliers for each group. 
```{r}
top_10_each_position <- position_sub %>%
                          summarize_each(funs(mean)) %>%
                          arrange(desc(avg_year)) %>%
                          head(10) %>%
                          select(position) %>%
                          inner_join(position_sub, by='position') %>% 
                          group_by(position) %>%
                          slice_max(order_by = avg_year, n = 10)

ggplot(top_10_each_position, aes(x=position, y=avg_year))+
      geom_boxplot() +
      theme(axis.text.x = element_text(angle=90, hjust = 1))

```
Even in the most-paid subset, **quarterback** has the obvious highest yearly pay.

#### age
To look at if age affect salary, we will first look at the most paid players in each position to see their age distribution.
```{r}
player_salary %>% group_by(position) %>% slice_max(order_by = avg_year, n = 10) %>%
  ggplot(aes(x=age)) + 
  geom_histogram(binwidth = 2,
                 color='black',
                 fill='white')+
   geom_density() +
   geom_density(alpha=.2, fill='#FF6666')+
   geom_vline(aes(xintercept=mean(age, na.rm=T)),
               color="red", linetype="dashed", size=1)
```
We can tell that the age of most top-paid players are concentrated around 26, and the overall distribution is slightly right-skewed.
```{r}
player_salary %>%
  ggplot(aes(x=age)) + 
  geom_histogram(binwidth = 2,
                 color='black',
                 fill='white')+
   # geom_density(alpha=.2, fill="#FF6666")+
   geom_vline(aes(xintercept=mean(age, na.rm=T)),
               color="red", linetype="dashed", size=1)
```
The overall age distribution has a very similar tendency to the top-paid one, the right-skewness is easier to identify. 30 seems like a division -- the number of frequency drops drasctially after 30. 

#### team
To look at the effect of team, we will look at both highest paid positions for each team and average pay for each team.  

```{r}
# Average
# head(player_salary)
# player_salary %>% group_by(team) %>% select(c('team','avg_year')) %>% summarize_each(funs(mean)) %>%
#   ggplot(aes(x=team))+geom_boxplot() + theme(axis.text.x = element_text(angle=90, hjust = 1))

p <-player_salary[,c('team','avg_year')] %>% group_by(team) %>%
      summarize_each(funs(mean)) %>%
      arrange(desc(avg_year)) %>%
head(32) %>%
select(team) %>%
inner_join(player_salary[,c('team','avg_year')], by='team') %>%
group_by(team)%>%
ggplot(aes(x=team, y=avg_year))+
geom_boxplot(aes(fill=avg_year))+
theme(axis.text.x = element_text(angle=90, hjust = 1))
p + stat_summary(fun=mean, color='darkred', geom='point') #red dot identifies the mean value
```
There are some fluctuations, zoom in to identify the tendency.
```{r}
p + stat_summary(fun=mean, color='darkred', geom='point') +
    coord_trans(x = "identity", y = "identity", xlim = NULL, ylim = c(0,4e+6))
```
Unlike the position, there is not an outstanding best-paid team, rather, there are few less-paid teams, so let's look at the bar plot.
```{r}
less_team <- player_salary[,c('team','avg_year')] %>% 
      group_by(team) %>%
      summarize_each(funs(mean)) %>%
      arrange(avg_year) %>%
      head(16) %>%
      select(team) %>%
      inner_join(player_salary[,c('team','avg_year')], by='team') %>%
      group_by(team) %>% 
      summarize_each(funs(mean)) %>% 
      mutate(Rate='Lo')
hi_team <- player_salary[,c('team','avg_year')] %>% 
      group_by(team) %>%
      summarize_each(funs(mean)) %>%
      arrange(desc(avg_year)) %>%
      head(16) %>%
      select(team) %>%
      inner_join(player_salary[,c('team','avg_year')], by='team') %>%
      group_by(team) %>% 
      arrange(desc(avg_year)) %>%  
      summarize_each(funs(mean)) %>%
      mutate(Rate='Hi')
hi_low_team <- rbind(hi_team, less_team)

ggplot(hi_low_team, aes(x=team, y=avg_year))+
  theme(axis.text.x = element_text(angle=90, hjust = 1))+
  geom_bar(stat = 'identity')
  # facet_grid(. ~ Rate)
```
The *Cowboy*, *Rams* obviously has lower average salary than the other teams. We know that *headquarter* is the position to have higher salary, so we want to know if these teams are paying lower salary depends on positions.
```{r}
all_team <- player_salary[,c('avg_year', 'position')]  %>% 
                group_by(position) %>%
                summarise_each(funs(mean))%>%
                mutate(Distr = 'All')

Cowboy_team <- 
            player_salary[player_salary$team=='Cowboys' ,c('avg_year', 'position')]%>% 
                group_by(position) %>%
                summarise_each(funs(mean))%>%
                mutate(Distr ='Cowboy')

Rams_team <- player_salary[player_salary$team=='Rams' ,c('avg_year', 'position')] %>% 
                group_by(position) %>%
                summarise_each(funs(mean))%>%
                mutate(Distr ='Rams') 

position_team <- rbind(all_team, Cowboy_team, Rams_team)
ggplot(position_team, aes(x=position, y=avg_year))+
  theme(axis.text.x = element_text(angle=90, hjust = 1))+
  geom_bar(stat = 'identity')+
  facet_grid(Distr ~ .)
```
Interestingly, neither of these follow the overall trend, but they both follow the same pattern: paying more to *4-3 defensive-tackle* and *left-tackle*. 


## NanNan
```{r}
head(football_salaries)
football <- football_salaries
football <- na.omit(football)
a <- c(0,323,2020,1)
football <- football[football$age!=0,]
football <- football[football$age!=323,]
football <- football[football$age!=2020,]
football <- football[football$age!=1,]

library(tidyverse)
# number of players of each position
frequency <- football %>% group_by(football$position)%>%
  summarise(n = n())
frequency <- as.data.frame(frequency)
x
frequency[order(frequency$n,decreasing = T),]
```
defensive-back is the position that has the most players across all positions
offensive-line is the position that has the second most players.

```{r}
# average salary of each team
teamSa <- football %>% group_by(football$team) %>%
  summarise(avg <- mean(total_value))
teamSa <- as.data.frame(teamSa)
teamSa[order(teamSa$`avg <- mean(total_value)`,decreasing=T),]
```
players in Vikings make the most across all teams.

```{r}
# average salary of each age group
unique(football$age)
ageSa <- football %>% group_by(football$age) %>%
  summarise(avg <- mean(total_value))
ageSa <- as.data.frame(ageSa)
ageSa[order(ageSa$`avg <- mean(total_value)`,decreasing=T),]
```
players older than 30 could make much more than those younger than 30

```{r}
# average salary of each position
posSa <- football %>% group_by(football$position) %>%
  summarise(avg <- mean(total_value))
posSa <- as.data.frame(posSa)
posSa[order(posSa$`avg <- mean(total_value)`,decreasing=T),]
```
quarterback is the position that makes the most money across all positions

```{r}
# age and total value correlation
cor(football$total_value,football$age)
```
It shows that age is positively correlated with salary but it's not strong correlation

```{r}
# is total value positively correlated with fully guaranteed?
cor(football$total_value,football$fully_guaranteed)
```
It shows that total_value and fully_guaranteed are highly correlated.
