---
title: "Football Analysis Report"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
# Libraries used
library(dplyr)
library(tidyverse)
library(tidyr)
```

## Part A: Prepare the data for analysis

### 1, Read in data

```{r echo=F}
#install package 'xlsx' if not installed yet
# 
# if (!"xlsx" %in% installed.packages()) install.packages('xlsx')
# 
# library('xlsx')
# 
# #read in data sheets
# sheet.index <- 1:3
# data.list <- list()
# for(i in sheet.index)
#   data.list[[i]]<-read.xlsx('football_salaries.xls', sheetIndex = i)
# 
# # Rename the datasets as we wanted
# player_salary <- data.list[[1]]
# coach_salary <- data.list[[2]]
# coach_record <- data.list[[3]]
```

Read from different files

```{r}
player_salary <- read.csv('player_salary.csv',
                    header = TRUE,
                    sep = ',')
coach_salary <- read.csv('coach_salary.csv',
                    header = TRUE,
                    sep = ',')
coach_record <- read.csv('coach_record.csv',
                    header = TRUE,
                    sep = ',')
```

### 2, Cleaning data

The following part describes how we clean the data. We hide the code to
avoid unreasonable length of this report.

#### 2.1 General information of the datasets

We first take a brief look at the datasets of their descriptive
statistics.

```{r}
summary(player_salary)
summary(coach_salary)
summary(coach_record)
```

Check if there are Na values.

```{r}
# Check if there are NA values
# for (i in sheet.index){
#   print(data.list[[i]] %>% is.na() %>% sum())
# }
print(player_salary%>% is.na() %>% sum())
print(coach_salary%>% is.na() %>% sum())
print(coach_record%>% is.na() %>% sum())
```

We see there are some cleaning should be done before we can use these
datasets:\
1. Take care of the missing value in **coach_salaries**\
2. Change the column names in **coach_record**\
3. For consistency, change salary-related variable types to double in
**coach_salary** and **coach_record**.\
4. Separate variables if needed.

##### 2.1.1. missing values

Delete some unreasonable and missing values.

```{r}
# replicate the original model for cleaning
player_salary_ <- player_salary
coach_salary_ <- coach_salary
coach_record_ <- coach_record

# Finding the missing value, and filled in with internet searching result
coach_salary_ <- coach_salary
coach_salary_[c(which(is.na(coach_salary)==TRUE, arr.ind = TRUE)[,1]),5] <- '2.5 million'
```

##### 2.1.2 Fix column names

change all the column names into underline format.

```{r}
colnames(coach_record_) <- c(coach_record_[1,])
coach_record_ <- coach_record_[-1,]
rownames(coach_record_) <- c(1:nrow(coach_record_))

colnames(coach_salary_) <- c(colnames(coach_salary_)[1:3], 'Annual_salary', 'Net_worth', colnames(coach_salary_)[6])
```

##### 2.1.3 change the variable type

2.1.3.1. **coach_salary**\
\* *Annual_salary* \* This is changed into numerical

```{r}
# for (i in 1:nrow(coach_salary_)){
#   coach_salary_[i, 'Annual_salary'] <- gsub('\\$', '', coach_salary_[i, 'Annual_salary'])
#   coach_salary_[i, 'Annual_salary'] <- gsub(' ' , '', coach_salary_[i, 'Annual_salary'])
# }
```

```{r}
no_mill <- c(which(!agrepl('million', coach_salary_$Annual_salary, ignore.case = T)))

# paste 'million' to uniform these cells
for (i in no_mill)
  coach_salary_[i,'Annual_salary'] <- paste(coach_salary_[i,'Annual_salary'], 'million', collapse = ' ')

# coach_salary_
```

```{r}
coach_salary_ <- separate(coach_salary_, Annual_salary, c('income', NA), ' ')

# coach_salary_
```

```{r}
# Keep numeric parts and change variable type to numeric
for (i in 1:nrow(coach_salary_))
  coach_salary_[i,'income'] <-
substr(coach_salary_[i, 'income'], 2, nchar(coach_salary_[i,'income']))

# coach_salary_
```

```{r}
coach_salary_[23,'income'] <- '0.1'
coach_salary_[24,'income'] <- '4'
coach_salary_[26,'income'] <- '3.025'
coach_salary_[30,'income'] <- '3.5'
coach_salary_[31,'income'] <- '0.5'
# coach_salary_
```

```{r}
coach_salary_ <- mutate(coach_salary_, 'Annual_salary'=as.numeric(coach_salary_$income)*1000000)
# coach_salary_
```

```{r echo=F}
# # get the index of the cells with out 'million'
# no_mill <- c(which(!agrepl('million', coach_salary_$Annual_salary, ignore.case = T)))
# 
# # paste 'million' to uniform these cells
# for (i in no_mill)
#   coach_salary_[i,'Annual_salary'] <- paste(coach_salary_[i,'Annual_salary'], 'million', collapse = ' ')
# 
# # separate the numeric parts and drop million
# coach_salary_ <- separate(coach_salary_, Annual_salary, c('income', NA), ' ')
# 
# # Keep numeric parts and change variable type to numeric
# for (i in 1:nrow(coach_salary_))
#   coach_salary_[i,'income'] <-
# substr(coach_salary_[i, 'income'], 2, nchar(coach_salary_[i,'income']))
# 
# # numerize the variable
# for (i in 1:nrow(coach_salary_)){
#   if (i %in% no_mill)
#     coach_salary_[i, 'income'] <- as.double(gsub(',', '', coach_salary_[i,'income']))
#   if (i==24)
#       coach_salary_[i, 'income'] <- as.double(substr(coach_salary_[i, 'income'], 1, (nchar(coach_salary_[i,'income'])-1)))
#   else
#     coach_salary_[i, 'income'] <- as.double(coach_salary_[i, 'income'])*1000000
# }
# coach_salary_ <- rename(coach_salary_, 'Annual_salary'='income')
```

-   *Net_salary* \* This is also changed to numeric variable, there were
    missing piece warning, which is waht we want to get rid of in the
    str type, so just ignore it.

```{r, warning=FALSE}
# Remove dollar sign and trim
for (i in 1:nrow(coach_salary_)){
  coach_salary_[i, 'Net_worth'] <- gsub('\\$', '', coach_salary_[i, 'Net_worth'])
  coach_salary_[i, 'Net_worth'] <- gsub(' ' , '', coach_salary_[i, 'Net_worth'])
}

coach_salary_ <- separate(coach_salary_, Net_worth, c('net', NA), 'million')

coach_salary_$net <- as.double(coach_salary_$net)*1000000

coach_salary_ <- rename(coach_salary_, 'Net_worth'='net')
```

2.1.3.2 **coach_record** Change this to double type

```{r}
for (i in 4:ncol(coach_record_))
  coach_record_[,i] <- as.double(coach_record_[,i])
```

##### 2.1.4. Separate variables

-   **player_salary** Seperate agency and end time. There is warning
    saying missing value which is taken care of in the next part.

```{r}

player_salary_ <- player_salary %>% separate(free_agency, c('agency_time', 'type'),' ')
# There's warning of NA's in the dataset

# Extract rows with NA values
# player_salary_[c(which(is.na(player_salary_)==TRUE, arr.ind = TRUE)[,1]),]

# These are not top players
# the number of missing data is not very large, so we just drop them
player_salary_ <- player_salary_[-c(which(is.na(player_salary_)==TRUE, arr.ind = TRUE)[,1]) ,]
rownames(player_salary_) <- c(1:nrow(player_salary_))

player_salary_$agency_time <- as.numeric(player_salary_$agency_time)
```

#### 2.1.5, Fix abnormal data

Replace age larger than 50 or smaller than 15 with mean age of the rest
age data.

```{r, include=FALSE}
age_mean <- player_salary_ %>% filter(age>15) %>% filter(age<50) %>% select(age)%>% summarize_each(funs(mean))

player_salary_[player_salary_$age < 15, 'age'] <- age_mean
player_salary_[player_salary_$age > 50, 'age'] <- age_mean
```

There are three 0 values in *avg_year*, so we just delete them.

```{r}
player_salary_ <- player_salary_[player_salary_$avg_year != 0,]
```

#### 2.1.6, Repeated variable

There were few positions in **player_salary** are the same, merge them
as needed: (1) **3-4-defensive-end** and **4-3-denfensive-end**\
(2) **3-4-denfensive-tackle** and **4-3-defensive-tackle** (both belong
to **defensive-line**)\
(3) **linebacker**, **traditional-linebacker**, and
**insider-linebacker**\
(4) **3-4-outside-linebacker** and **4-3-outside-linebacker** There are
some positions belongs to the other, they are:\
**cornerback** and **safty** belong to **defensive-back**\
**center** belong to **offensive-line**\
**3-4-denfensive-tackle** and **4-3-defensive-tackle** (both belong to
**defensive-line**)

```{r}
player_salary_[player_salary_$position=='3-4-defensive-end','position'] <- '4-3-defensive-end'

player_salary_[player_salary_$position=="3-4-defensive-tackle",'position'] <- "4-3-defensive-tackle"

player_salary_[player_salary_$position=='traditional-linebacker','position'] <- "linebacker"
player_salary_[player_salary_$position=="inside-linebacker",'position'] <- "linebacker"

player_salary_[player_salary_$position=='3-4-outside-linebacker','position'] <- '4-3-outside-linebacker'
```

Here is the data we eventually have:

```{r}
player_salary <- player_salary_
coach_record <- coach_record_
coach_salary <- coach_salary_
head(player_salary)
head(coach_record)
head(coach_salary)
```


## Part B: Data analysis
### Introduction
In this project, we have three datasets about the basic information of the players and coach, including their salary, value, age, performances and so on. We will do analysis on how their income vary in terms of their age, position, winning rate and team. 
### 1 Analysis of Total value

We want to know how the *total_value* varies by age, position, and team.
We'll also do computations to showsome correlation.

#### 1.1 average total value of each position

```{r}
posSa <- player_salary_ %>% group_by(player_salary_$position) %>%
  summarise(avg <- mean(total_value))
posSa <- as.data.frame(posSa)
posSa[order(posSa$`avg <- mean(total_value)`,decreasing=T),]
```

This slide shows that quarterback position has the highest average
salary. Positions like quarterback, edge-rusher, tackle, and
4-3-defensive-end positions make over 10 millions.

#### 1.2 number of players of each position

```{r}
frequency <- player_salary_ %>% group_by(player_salary_$position)%>%
  summarise(n = n())
frequency <- as.data.frame(frequency)
frequency[order(frequency$n,decreasing = T),]
```

There are 837 linebackers a part of defense which is responsible to stop
both the run and pass. Their position is in the middle of the defense.
From this slide, we can see that teams tend to have more players in
defense. Wide-receiver is the position in the offense that has the
largest number of players.

#### 1.3 average salary of each team

```{r}
teamSa <- player_salary_  %>% group_by(player_salary_ $team) %>%
  summarise(avg <- mean(total_value))
teamSa <- as.data.frame(teamSa)
teamSa[order(teamSa$`avg <- mean(total_value)`,decreasing=T),]
```

This shows that the team Vikings earns the most in terms of the total
value which is the amount of money on the contract.

#### 1.4 average total value of each age group

```{r}
unique(player_salary$age)
ageSa <- player_salary_  %>% group_by(player_salary_ $age) %>%
  summarise(avg <- mean(total_value))
ageSa <- as.data.frame(ageSa)
ageSa[order(ageSa$`avg <- mean(total_value)`,decreasing=T),]
```

From this plot, we can see that players around 40 years old earn the
most compared to younger players.

#### 1.5 age and total value correlation

```{r}
cor(player_salary_ $total_value,player_salary_ $age)
corAge <- ggplot(data=player_salary_ ,aes(x=age,y=total_value))+xlim(0, 70)+geom_point(color ="darkgreen")
corAge
```

It shows that age is positively correlated with salary but it's not
strong correlation; From the scatter plot, we could see that players
with age around 30 make more money compared to other age groups, which
is consistent with finding from previous part.

#### 1.6 is total value positively correlated with fully guaranteed?

```{r}
cor(player_salary_ $total_value,player_salary_ $fully_guaranteed)
corValue <- ggplot(data=player_salary_ ,aes(x=total_value,y=fully_guaranteed))+
  geom_point(color ="darkgreen")
corValue
```

It shows that total_value and fully_guaranteed are highly correlated.

### 2 What are the variables related to high salary of players?

We discuss the how **total_value** varies in the part 1, now we are
interested in how the player are paid year by year, which is
**avg_year** so they have such huge accumulative **tptal_value**. We
propose the following:

#### 2.1 position

Position is the most intuitively reason we would associate with income
-- there are positions more important in a game, they are supposed to
make more money.

##### 2.1.1 Average highest

I take the average over the annual salary after grouping them by
position.

```{r}
# unique(player_salary$position)#24 positions in the data

# Select columns of position, avg_year, player to find positions with the highest salary

position_sub<- player_salary[c('position','avg_year')]%>%
  group_by(position)

p <- position_sub %>%
      summarize_each(funs(mean)) %>%
      arrange(desc(avg_year)) %>%
      head(10) %>%
      select(position) %>%
      inner_join(position_sub, by='position') %>% 
      group_by(position) %>%
      ggplot(aes(x=position, y=avg_year))+
      geom_boxplot(aes(fill=avg_year)) +
      theme(axis.text.x = element_text(angle=90, hjust = 1))
p + stat_summary(fun=mean, color='darkred', geom='point') #red dot identifies the mean value
```

If we look at the number of positions by `unique()`, we see there are 24
positions, but we are concentrating on the highest ones, so I keep the
top 10 average salary positions to display in the plot. Red dot in the
plot identifies the average value.\
Though the boxes are scaled by large value outliers, we cannot remove
those outliers, because they represent the extraordinarily high paid
players in each position, we'll take care of them later, Now I will zoom
in to show the tendency, of these average values.

```{r}
p + stat_summary(fun=mean, color='darkred', geom='point') +
    coord_trans(x = "identity", y = "identity", xlim = NULL, ylim = c(0,6e+6))

```

If we look at this zoomed in plot, we find the avarage annual salary of
**quarterback** is getting close to 6 million, which could be twice of
the number of 4-3 outside linebacker, but there are still 14 positions
not showed here, so **quarterback** is indeed a highly-paid position.

If we take a closer look at the boxplot, we find all the medians are
close to the bottom of the box and much lower than the red dots, which
means high-paid players extensively raise the mean value, so we would
like to look at those extraordinarily high paid players, i.e. the
outliers mentioned before.

##### 2.1.2 Top highest in each position

Here we extract top 10 highest paid players in top 10 high-paid
positions, and plot the average annual salary versus position,

```{r}
top_10_each_position <- position_sub %>%
                          summarize_each(funs(mean)) %>%
                          arrange(desc(avg_year)) %>%
                          head(10) %>%
                          select(position) %>%
                          inner_join(position_sub, by='position') %>% 
                          group_by(position) %>%
                          slice_max(order_by = avg_year, n = 10)

ggplot(top_10_each_position, aes(x=position, y=avg_year))+
      geom_boxplot() +
      theme(axis.text.x = element_text(angle=90, hjust = 1))

```

This plots shows that even in the most-paid subset, **quarterback** has
the obvious highest yearly pay. If we look at the overall trend of each
position, they follow a similar trend we had in the previous box plot.

#### 2.2 age

Another factor matters to salary could be age. It is conceivable that a
player at younger age tend to be more energetic and active while older
players are more experience at team-work and techniques, so there could
be an optimum age preferred by clubs and willing to pay more.

##### 2.2.1 Top player age distribution

To look at if age affect salary, we will first look at the most paid
players in each position to see their age distribution.

```{r}
player_salary %>% group_by(position) %>% slice_max(order_by = avg_year, n = 10) %>%
  ggplot(aes(x=age)) + 
  geom_histogram(binwidth = 2,
                 color='black',
                 fill='white')+
   geom_density() +
   geom_density(alpha=.2, fill='#FF6666')+
   geom_vline(aes(xintercept=mean(age, na.rm=T)),
               color="red", linetype="dashed", size=1) #red dash is the mean
```

This plot has bin width of 2, and the red dashed line marks the mean of
the distribution. We found most top-paid players have age between 25 and
31, and the mean is 29, very few of them are over 40, so this
distribution is slightly right-skewed.

##### 2.2.1 All players age distribution

We also interested in age distribution of all players, so we plot this:

```{r}
player_salary %>%
  ggplot(aes(x=age)) + 
  geom_histogram(binwidth = 2,
                 color='black',
                 fill='white')+
   # geom_density(alpha=.2, fill="#FF6666")+
   geom_vline(aes(xintercept=mean(age, na.rm=T)),
               color="red", linetype="dashed", size=1)
```

The overall age distribution has a similar tendency to the top-paid one:
right tail, but much heavy this time. This reflects that younger players
are preferred in this game. Another evidence verifies younger preference
is the mean, in this plot, the mean decreases to 26, which was 19 in the
previous plot. This young-age preference is understandable, because
aging and experience don't guarantee for good performance.

#### 2.3 team

Just like some companies are willing to pay more to their employees,
annual salary can vary by teams.

##### 2.3.1 Average salary versus team

We take average of each team and plot the following:

```{r}
# Average
# head(player_salary)
# player_salary %>% group_by(team) %>% select(c('team','avg_year')) %>% summarize_each(funs(mean)) %>%
#   ggplot(aes(x=team))+geom_boxplot() + theme(axis.text.x = element_text(angle=90, hjust = 1))

p <-player_salary[,c('team','avg_year')] %>% group_by(team) %>%
      summarize_each(funs(mean)) %>%
      arrange(desc(avg_year)) %>%
head(32) %>%
select(team) %>%
inner_join(player_salary[,c('team','avg_year')], by='team') %>%
group_by(team)%>%
ggplot(aes(x=team, y=avg_year))+
geom_boxplot(aes(fill=avg_year))+
theme(axis.text.x = element_text(angle=90, hjust = 1))
p + stat_summary(fun=mean, color='darkred', geom='point') #red dot identifies the mean value
```

As what happened in position, there are outliers in each team. There
seems a fluctuation, zoom in to identify the tendency.

```{r}
p + stat_summary(fun=mean, color='darkred', geom='point') +
    coord_trans(x = "identity", y = "identity", xlim = NULL, ylim = c(0,4e+6))
```

Unlike the position, there is not an outstanding best-paid team, rather,
there are few less-paid teams, so let's look at the bar plot.

```{r}
less_team <- player_salary[,c('team','avg_year')] %>% 
      group_by(team) %>%
      summarize_each(funs(mean)) %>%
      arrange(avg_year) %>%
      head(16) %>%
      select(team) %>%
      inner_join(player_salary[,c('team','avg_year')], by='team') %>%
      group_by(team) %>% 
      summarize_each(funs(mean)) %>% 
      mutate(Rate='Lo')
hi_team <- player_salary[,c('team','avg_year')] %>% 
      group_by(team) %>%
      summarize_each(funs(mean)) %>%
      arrange(desc(avg_year)) %>%
      head(16) %>%
      select(team) %>%
      inner_join(player_salary[,c('team','avg_year')], by='team') %>%
      group_by(team) %>% 
      arrange(desc(avg_year)) %>%  
      summarize_each(funs(mean)) %>%
      mutate(Rate='Hi')
hi_low_team <- rbind(hi_team, less_team)

ggplot(hi_low_team, aes(x=team, y=avg_year))+
  theme(axis.text.x = element_text(angle=90, hjust = 1))+
  geom_bar(stat = 'identity')
  # facet_grid(. ~ Rate)
```

In this bar plot, we can see the average annual of each team fluctuates
between 2 million and 2.5 million without outstanding high value. But
there are two teams: *Dolphins*, *Ravens* are obviously have lower
values than the other teams. We know that *headquarter* is the position
to have higher salary, so we want to know if these teams are paying
lower salary depends on positions.

Here we plot the average annual salary distributions by positions. To
compare the how are the position investing strategy of *Dolphins* and
*Ravens* are different from the overall trend, we put the plots
together.

```{r}
all_team <- player_salary[,c('avg_year', 'position')]  %>% 
                group_by(position) %>%
                summarise_each(funs(mean))%>%
                mutate(Distr = 'All')

Dolphins_team <- 
            player_salary[player_salary$team=='Dolphins' ,c('avg_year', 'position')]%>% 
                group_by(position) %>%
                summarise_each(funs(mean))%>%
                mutate(Distr ='Dolphins')

Ravens_team <- player_salary[player_salary$team=='Ravens' ,c('avg_year', 'position')] %>% 
                group_by(position) %>%
                summarise_each(funs(mean))%>%
                mutate(Distr ='Ravens') 

position_team <- rbind(all_team, Dolphins_team, Ravens_team)
ggplot(position_team, aes(x=position, y=avg_year))+
  theme(axis.text.x = element_text(angle=90, hjust = 1))+
  geom_bar(stat = 'identity')+
  facet_grid(Distr ~ .)
```

These three plots are different in their own way, but one noteworthy
position is *Safty*. Both *Dolphins* and *Ravens* are willing to pay a
lot to this position. But if we look at the top most plot, that's not
the case. It's probably that these two teams have good defense
performance in the games.

### 4. How many players will be free agency by 2022?

The graph below shows the number of free agency players per year. As
explained earlier, free agency means that these players will be free to
negotiate with other teams by the year that their contract ends. As we
can see in the graph, there will be almost 5,000 free agency players by
2022, taking in count from 2017 to 2022. Also, the figure shows a trend
of decreasing the number of free agency players from 2020 to 2025. This
trend might be due to many players ending their contract to start a new
one.

```{r}
# Cheking how many player will be free agency by 2022
df_free <- player_salary %>% dplyr::filter(agency_time >= 2022)
nrow(df_free)

# Plotting
df_agency <- player_salary
df_agency$agency_time <- as.factor(df_agency$agency_time)

ggplot(df_agency, aes(x= agency_time, color="agency_time")) +
    geom_bar(stat = "count", fill="white") +
    labs(x = "Free Agency") +
  #ylim(0, 2500) +
  scale_y_continuous(breaks = seq(0, 2500, by = 250))
  ggtitle("Number of players by free agency years")
```

### 5. What are the 20 MVP players based on their earns?

The list below presents the 20 MVP, which stand for the 20 Most Valuable
Players from the NFL league. For that calculation, we considered values
from the dataset such as average per year (avg_year), total_guaranteed,
and fully_guaranteed. As we can see, most of the player's positions on
this list are a quarterback. The top three were Matt Ryan (Falcons),
Russel Wilson (Seahawks), and Aaron Rodgers (Packers).

```{r}
# Creating column with MVP value 
df_MVP <- player_salary %>% 
  rowwise() %>% 
  mutate(MVP_value = sum(avg_year, total_guaranteed, fully_guaranteed))

# Removing duplicated players'name
df_MVP <- df_MVP %>% distinct(player, .keep_all= TRUE)
# Organizing it by sorting as ascending values
df_MVP <- df_MVP[order(df_MVP$MVP_value),]
# Extracting the 20 highest MVP_values
df_20 <- tail(df_MVP, n= 20)
# The 20 most valuable players (MVP) are:
df_20$player
# Which position has more MVP's?
df_20$position <- as.factor(df_20$position) # making position as factor

# Removing duplicated players'name
df_MVP <- df_MVP %>% distinct(player, .keep_all= TRUE)

# Organizing it by sorting as ascending values
df_MVP <- df_MVP[order(df_MVP$MVP_value),]

# Extracting the 20 highest MVP_values
df_20 <- tail(df_MVP, n= 20)

# The 20 most valuable players (MVP) are:
df_20$player

# Which position has more MVP's?
df_20$position <- as.factor(df_20$position) # making position as factor
```

As mentioned in the previous graph, the position with more MVPs based on
our dataset is the quarterback, followed by the defensive line,
4-3-outside linebacker, and finally wide-receiver. Since the
quarterbacks are responsible for creating plays for all the team, it
makes sense that they make more money.

```{r}
ggplot(df_20, aes(x= position, color= position, fill= position)) +
    geom_bar(stat = "count") +
    labs(x = "Position") +
  theme(legend.position = "none")+
  ggtitle("MVP's per position")
```

### 6. What teams have the highest fully guaranteed expenses?

The graph below shows the teams with higher fully guaranteed expenses,
which is the money that the teams will spend with players regardless of
what happens (injury, suspension, etc.). The data was different from
expected because neither of the teams that won the Super Bowl in the
last three years (New England Patriots, KC Chiefs, and Tampa Bays) has
more expenses due to higher investments. Therefore, more analysis is
needed for different assumptions.In this case, the Bears team has the
highest fully guaranteed expenses (\~4,2 M dollars), followed by Red
Skins (\~4,1 M dollars) and Browns (\~3,8 M dollars).

```{r}
# Grouping by teams
library(dplyr)
df_fully <- player_salary %>% group_by(team) %>% dplyr::summarize(fully_guaranteed_mean=(mean(fully_guaranteed)))

ggplot(player_salary, aes(x= reorder(team, +fully_guaranteed), y=fully_guaranteed, color= team, fill= team)) +
    geom_col() +
    labs(x = "Team", y= "Fully guaranteed expenses ($)") + 
  theme(legend.position = "none") +
  ggtitle("Teams with highest fully guaranteed expenses") + 
    coord_flip()
```

### 7. What is the relationship between coach salaries and their records?

Finally, the graph below shows some relationship between coach records
wins and their salaries, where we were trying to find a trend on that.
We hypothesize that the percentage of wins per number of games from
these coaches is highly related to how much money they make. However, as
we can see in the graph, we have no clear trend on this relationship. If
we don't consider an outlier, the coach salaries ranged from 2.5 million
to 10 million. Maybe after 60% of wins, some coaches are showing an
annual salary above 7.5 million. The coach that is making more money in
the league is Andy Reid from KC Chiefs, which makes sense because he was
in the last two Super Bowls and won one of them. Other than that, we
would have to analyze more data to make some conclusions, maybe even
looking at the number of losses, championships, etc.

```{r}
coach_record <- rename(coach_record, name = Name)
df_coach <- coach_record %>% merge(coach_salary, by = "name", all.x = TRUE)  %>% 
  na.omit()
```

```{r}
ggplot(df_coach, aes(x= Pct*100, y=Annual_salary, color=name)) +
    geom_point(size=4, alpha=0.6) +
    labs(x = "Wins (%)", y= "Salary ($)") + 
  #theme(legend.position = "none") +
  coord_trans(x= "identity", y= "identity", xlim= NULL,  ylim= c(0, 10e+6 )) 
```

## Part C Conclusion
In this project, we are interested in how are the income of football players and coach are affected by various factors.  
After comparing the **total_value**, we find the value of a player is related to their age, position and shows they are statistically correlated bu computing correlations.  
To explore more about how the players has those accumulated total value, we also did some analysis regarding their annual salary (**avg_year**) and factors have effect on those. We found younger age (25 or so) are likely to have higher salary.  
Though the position-wise invest strategy varies team by team, *quarterback* in general is a good choice for high salary, which can be attributed to the number quarterback players in top 20 MPV.       
Good performance ones earns more rule is true to players, but there is no clear evidence show the same case for coached -- the winning rate and coach salary doesn't show an obvious correlation.