---
title: "Stat 850 Project Description"
author: "Nannan Wang; Muxin Hua; Jose Cesario."
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
```

```{r load-sas-libraries, echo = F, message = F, warning = F}

library(SASmarkdown)
sas_enginesetup(sashtml=sashtml)

sasexe <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
sasopts <- "-nosplash -ls 75"

# Linux options (for grading, please leave this in!)
# sasexe <- "/usr/local/SASHome/SASFoundation/9.4/bin/sas_en"
# sasopts <- "-ls 75"
```

## Dataset

* Datasets
We have three datasets for this project, **player_salary**, **coach_salary**, and **coach record**.

* Links
  + **player_salary** https://www.kaggle.com/trolukovich/football-players-salaries?fbclid=IwAR0sLUs9Js9-CpRRI9y_3QJKPuMxelc6Na77423U4v3KVzVw3Vbo9iK2x4U
  + **coach_salary** https://sportscriber.com/money/nfl-coaches-salary/  
  + **coach record** https://www.footballdb.com/coaches/index.html?fbclid=IwAR0Um2j5_B3LaPOX8jg2hl_K_LSc-zlK1cA0iYxTvYtnPWCgPJ6B40yEKJY

* Variables
  + **player_salary**  
      *position*         : position played by the player   
      *player*           : name   
      *team*             : team the player in   
      *age*              : age     
      *total_value*      : total amount of money on the contract
      *ave_year*         : yearly averaged salary  
      *total_guaranteed* : players could get this amount of money after they serve several years(from the contract) for the team  
      *fully_guaranteed* : money guaranteed to get regardless of what happens    
      *free_agency*      : time when the contract of the player expires   
      </br>
  + **coach_salary**  
      *name*: name of the coach  
      *team*: team the coach works for  
      *division*: division of the team  
      *annual salary*: annual salary   
      *net worth*: Total sum of property   
      *position*: position of the coach  
      </br>
  + **coach record**      
      *name*: name of the coach  
      *Team*: team the coach works for  
      *seasons*: seasons the coach coaches   
      *Gms*: games Coached  
      *W*: games Won  
      *L*: games Lost  
      *T*: games Tied  
      *Yrs*: seasons as head coach  
      *post*: number of postseason appearances as head coach  
      *pct*: winning Percentage  
      *DIV*: division Championships  
      *CON*: conference Championships  
      *LG*: league Championships  
      *SB*: super Bowl Championships  
      </br>
* Not all our datasets are from the same year. If we look at the time range of contracts,  most of them are valid for 5 years, so we assume the situations don't change much in few years in the field, which justifies the conclusions from these datasets. 
      
## Potential Topics to Explore Using the Data Set

* What are the variables related to high salary of players?
  + position
  + age
  + team
  </br>
* Is there salary difference in terms of:
  + age
  + position
  </br>
* correlation exploration:
  + is age correlated with player's total value?
  + is total value positively associated with fully-guaranteed? 
  </br>
* What are the teams tend to pay more.  
  + Join with coach data to see if these teams tend to pay more to their coaches too.  
  + Join with coach record data to see if high salary related to coach career history.  
  </br>

* How many players will be free agency by 2022?
  + Check the relationship between free agency players and their age
  + What teams will have more free agency players upcoming?
  
* What are the 20 MVP players based on their earns?
  + Based on avg year, total guaranteed and fully guaranteed amounts

* What teams have the highest fully guaranteed expenses?
  + Join coach data to see which team have the most guaranteed expenses per year

* What is the relationship between coach salaries and their records?
  + Considering their number of wins, losses, ties, championships


## Group Members
Nannan Wang  
Muxin Hua   
Jose Cesario
