---
title: "Stat 850 Project Description"
author: "Firstname Lastname, Firstname Lastname, ..."
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(digits = 3)
```

```{r load-sas-libraries, echo = F, message = F, warning = F}

library(SASmarkdown)
sas_enginesetup(sashtml=sashtml)

sasexe <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
sasopts <- "-nosplash -ls 75"

# Linux options (for grading, please leave this in!)
# sasexe <- "/usr/local/SASHome/SASFoundation/9.4/bin/sas_en"
# sasopts <- "-ls 75"
```


## Instructions
  
Each member of your team should modify this document in some way and push their modifications to the repository in a separate commit. This will ensure that you have set your repository up so that all group members are working with the same repository. 

Note that you must compile your readme (this document) for it to properly display as part of your github repository. 

Once you have received feedback on your project proposal (via Canvas) you may alter this README so that it describes your final project instead of the project proposal. 

## Data Set
Provide a link to your dataset here, and a description of the variables in the data set.

## Potential Topics to Explore Using the Data Set
Describe avenues you might explore using the data. What are your goals? How do you plan to explore the data? You can provide some exploratory plots here if you have them available, otherwise, list what types of things you will be looking for.

## Group Members
List all of the project contributors here.

